From Freebase to Wikidata: The Great Migration

### 2015-10-14 paper.tex
PDFs for figures. Strict ACM style.
Wrote Section 4.3. Added another screenshot.

\documentclass{sig-alternate}

@@

\numberofauthors{3}
\author{
Thomas Pellissier Tanon,\titlenote{The author was an intern at Google when the majority of the work was done.}\ Denny Vrandečić\\
   \affaddr{Google, San Francisco, United States of America}\\   
   \email{\{thomaspt, vrandecic\}@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google, Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\alignauthor
Thomas Steiner\\
   \affaddr{Google, Hamburg, Germany}\\   
   \email{tsteiner@google.com}
}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-statement-obama.png}
\caption{Target Wikidata statement}
\label{fig:statement-obama}
\end{figure}

@@

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements for verification to the user
so that these statements can be added to the currently shown Wikidata item.
With just one click, the user can reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The code of the Primary Sources Tool is openly available
under the terms of the Apache~2.0 license (\url{https://github.com/google/primarysources}).
The tool is deployed as a~so-called gadget,
\emph{i.e.}, as an easily activatable add-on feature in Wikidata.
It is independent from Freebase and can easily be reused to import other datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~dataset filter in the front-end that allows users
to display only statements from a~selected dataset.
This flexibility is indeed much-wanted, and during the development period of the project
another researcher approached us in order to upload
a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available publicly at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
At the time of writing (October, 2015), the tool has been used by more than a~hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely with the Wikidata user interface.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/primary-sources.png}
\caption{Primary Sources Tool user interface}
\label{fig:primary-sources-tool}
\end{figure}

@@

In order to prepare the data for integration into the Primary Sources Tool,
we have created a set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code is available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping tool.
We first created a~small dataset of around 1M statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.
In the following, we progressively added more statements to the tool's database back-end
and made the opt-in process to use the tool in Wikidata easier for users.
Initially, in order to use the tool, users had to manually copy and paste a~line
in their individual \texttt{common.js} user profiles,
whereas after the conversion to a~Wikidata gadget,
the tool could simply be activated via an option in the user preferences.
Thereby, we were able to both slowly increase the number of users of the tool
and upscale the tool's database back-end without risking performance bottlenecks.
The process of adding data in small batches of 1M~to~2M statements per batch allowed us
to detect some potential performances issues in the tool and fix them
before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts easily
and correct minor issues with the data early on already in the back-end.

@@

\todo{Tom: write me}

@@

\begin{itemize}
  \setlength\itemsep{0em}
  \item 48M~topics
  \item 2,997M triples
  \item 442M facts owned by Google%
    \footnote{These facts are the remaining Freebase triples
      after filtering out data about IDs (\texttt{/type/object/id}),
      types (\texttt{/type/object/type}), labels, descriptions (\texttt{/type/object/name},
      \texttt{/common/topic/description}, \emph{etc.}), and triples that were removed
      due to their licensing (see \autoref{sec:licensing})}
  \item 68M~labels
\end{itemize}

@@

\begin{figure*}[!htbp]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/freebase-time-space.png}
        \caption{Freebase}
        \label{fig:freebase}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/wikidata-time-space.png}
        \caption{Wikidata}
        \label{fig:wikidata}
    \end{subfigure}
    \caption{Spatio-temporal distribution of Freebase and Wikidata items,
      the $x$ axis shows longitude degrees from 180°~E to 180°~W, the $y$ axis shows time
      from the year 0 to 2015, following a~power law ($y' = y^8$)}
    \label{fig:time-space}
\end{figure*}

>>>>

%\documentclass{sig-alternate}
\documentclass{acm_proc_article-sp}

@@

\numberofauthors{3}
\author{
Thomas Pellissier Tanon,\titlenote{The author was an intern at Google when the majority of the work was done.}\ Denny Vrandečić\\
   \affaddr{Google San Francisco, United States of America}\\   
   \email{\{thomaspt, vrandecic\}@google.com}
\alignauthor
Sebastian Schaffert\\
   \affaddr{Google Zurich, Switzerland}\\   
   \email{schaffert@google.com}
\alignauthor
Thomas Steiner\\
   \affaddr{Google Hamburg, Germany}\\   
   \email{tsteiner@google.com}
}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{img/wikidata-statement-obama.pdf}
\caption{Target Wikidata statement}
\label{fig:statement-obama}
\end{figure}

@@

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements for verification to the user
so that these statements can be added to the currently shown Wikidata item.
With just one click, the user can reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The code of the Primary Sources Tool is openly available
under the terms of the Apache~2.0 license (\url{https://github.com/google/primarysources}).
The tool is deployed as a~so-called gadget,
\emph{i.e.}, as an easily activatable add-on feature in Wikidata.
It is independent from Freebase and can easily be reused to import other datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~dataset filter in the front-end that allows users
to display only statements from a~selected dataset.
This flexibility is indeed much-wanted, and during the development period of the project
another researcher approached us in order to upload
a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available publicly at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
At the time of writing (October, 2015), the tool has been used by more than a~hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows two screenshots
of the tool that integrates nicely with the Wikidata user interface.

\begin{figure}[!htbp]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/primary-sources.pdf}
        \caption{Incoming references}
        \label{fig:barack-obama}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-cause-of-death.pdf}
        \caption{Incoming claims and references}
        \label{fig:cause-of-death}
    \end{subfigure}
    \caption{Primary Sources Tool user interface, incoming data from Freebase shown in light blue}
    \label{fig:primary-sources-tool}
\end{figure}

@@

In order to prepare the data for integration into the Primary Sources Tool,
we have created a set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code is available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping tool.
We first created a~small dataset of around 1M statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.
In the following, we progressively added more statements to the tool's database back-end.
Thereby, we were able to slowly upscale the tool's database back-end
without risking back-end performance bottlenecks.
The process of adding data in small batches of 1M~to~2M statements per batch
further allowed us to detect some potential issues in the front-end
and fix them before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts easily
and correct minor issues with the data early on already in the back-end.
Upon approval or rejection of either a~claim or a~reference,
the Wikidata item page in question reloads in order to reflect
the potentially changed state of the migration, as,
in case of incoming claims with references, freshly accepted claims 
can require incoming references to be attached elsewhere in the hierarchy.

@@

For the front-end, our objective was to make the migration process
as close to the natural Wikidata editing flow as possible.
Ideally, the user should barely notice the different underlying datasources
when navigating Wikidata.
To achieve this, we have in a~first step created a~Wikidata \emph{user script}.
Wikidata user scripts are part of the Wikidata tool chain%
\footnote{Wikidata tool chain:
\url{https://www.wikidata.org/wiki/Wikidata:Tools}}
and are created by users, but unlike gadgets---to be explained in the following---%
do not appear in a~user's preferences.
Instead, one has to add a~line of code into ones \texttt{common.js} file
in order to activate and use the script.
User scripts can be created by anyone and might not always be stable.
Once a~user script has matured, it can be converted into a~\emph{gadget}.
Gadgets are scripts that are likewise created by users,
but which can be simply enabled in user preferences under the section ``Gadgets''.
They can only be edited by administrators and are assumed to be stable.
With our front-end, we went through this maturation steps
and started with a~user script called \texttt{freebase2wikidata.js}%
\footnote{Initial user script:
\url{https://www.wikidata.org/wiki/User:Tomayac/freebase2wikidata.js}}
that we later converted into a~gadget.%
\footnote{Final gadget: \url{https://www.wikidata.org/wiki/MediaWiki:Gadget-PrimarySources.js}}
\autoref{fig:primary-sources-tool} illustrates the user interface in action.
Existing Wikidata statements are shown as usual,
Freebase facts that could potentially be imported into Wikidata are shown in light blue
with instead of an ``edit'' button two new buttons ``approve reference'' and ``reject reference''
(\autoref{fig:barack-obama}) or ``approve claim'' and ``reject claim''
(\autoref{fig:cause-of-death}) respectively.

@@

\begin{itemize}
  \setlength\itemsep{0em}
  \item 48M~topics
  \item 2,997M triples
  \item 442M facts owned by Google%
    \footnote{That is, the remaining Freebase triples
      after filtering out data about IDs (\texttt{/type/object/id}),
      types (\texttt{/type/object/type}), labels, descriptions (\texttt{/type/object/name},
      \texttt{/common/topic/description}, \emph{etc.}), and triples removed
      for licensing (see \autoref{sec:licensing})}
  \item 68M~labels
\end{itemize}

@@

\begin{figure*}[!htbp]
    \centering
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/freebase-time-space.png}
        \caption{Freebase}
        \label{fig:freebase}
    \end{subfigure}
    \begin{subfigure}[b]{1.0\columnwidth}
        \includegraphics[width=\textwidth]{img/wikidata-time-space.png}
        \caption{Wikidata}
        \label{fig:wikidata}
    \end{subfigure}
    \caption{Spatio-temporal distribution of Freebase and Wikidata items,
      the $x$ axis shows longitude degrees from 180°~E to 180°~W, the $y$ axis shows time
      from the year 0 to 2015, following a~power law ($y' = y^8$) \todo{Create PDF versions of both diagrams}}
    \label{fig:time-space}
\end{figure*}
